---
title: "simulation-regression-method"
output:
  html_document:
    theme: flatly
    highlight: tango
---
```{r include=FALSE}
library(tidyverse)
```

## simulate dataset for regression

```{r}
#generate x1
X1 = rnorm(100,mean=3,sd=1)
#generate x2
X2 = rnorm(100,mean=2,sd=2)
#generate x3(highly correlated with x2)
X3 = 0.1*X2^2+1
#generate x4
X4 =rnorm(100,mean=3,sd=1.5)
#generate y
eps = rnorm(100,mean=0,sd=2)
Y = 2+1.5*X1+2*X3+0.75*X4+eps
#X matrix
data=data.frame(Y,X1,X2,X3,X4)
```

## Model1-Two-stage regression

```{r pressure, echo=FALSE}
#multivariate regression
reg = lm(Y~X1+X2+X3+X4,data)
summary(reg)
```

```{r}
reg_step1 = lm(Y~X1+X2+X3,data)
data = data%>%mutate(residual_Y = Y- predict(reg_step1))
reg_step2 = lm(X4~X1+X2+X3,data)
data = data%>%mutate(residual_X4 = X4- predict(reg_step2))
reg_step3 = lm(residual_Y~residual_X4,data)
summary(reg_step3)
```

Comparing the result, we can see the coefficient of the fourth parameter(X4) calculated by two-stage regression is the same as the result in multivariate regression.


## Model2: Successive orthogonalization



```{r}
##by definition
Z=matrix(nrow=100,ncol=5)
##can set any start point to find orthogonalized vectors
Z[,1]=rep(1,100)
R=matrix(nrow=100,ncol=4)
cof0=Z[,1]%*%X1/(Z[,1]%*%Z[,1])
Z[,2] = X1-cof0%*%Z[,1]
```

```{r}
proj <- function(y, X, list=FALSE) {
  if (is.vector(y)) y <- matrix(y, ncol=1)
  if (is.vector(X)) X <- matrix(X, ncol=1)
  XPX <- crossprod(X)
  P <- X %*% MASS::ginv(XPX) %*% t(X)
  if (!list) return(c(P %*% y))
  else return(list(y=c(P %*% y), P=P))
}
Z[,3] <- X2 - proj(X2, Z[,1])-proj(X2,Z[,2])
Z[,4] <- X3 - proj(X3, Z[,1]) - proj(X3, Z[,2])-proj(X3,Z[,3])
Z[,5] <- X4 - proj(X4, Z[,1]) - proj(X4, Z[,2]) - proj(X4, Z[,3])-proj(X4,Z[,4])
summary(lm(Y~Z[,5]))
```


Comparing the result, we can see the coefficient of the fourth parameter(X4) is the same as the result in multivariate regression.